---
title: "hw02-sandeep-tiwari"
author: "Sandeep Tiwari"
date: "9/29/2017"
output: github_document
---

# 2. Import the data in R
```{r}
#install.packages("readr")
library("readr")
library("dplyr")
library("ggplot2")
```

```{r}
dat <- read.csv("data/nba2017-player-statistics.csv", 
                colClasses = c("Player"="character", "Team"="character", "Experience"="character", "Position"="factor", "Salary"="double"))
```

```{r}
nba <- read_csv("data/nba2017-player-statistics.csv",
                col_types = cols(Player = col_character(), Team = col_character(), Experience = col_integer(), Position = col_factor(levels = c("PG", "SG", "SF", "PF", "C")), Salary = col_double()))
```

```{r}
str(dat)
str(nba)
```

# 3. Right after importing the data
### Replacing all occurrences of "R" in Experience vector with 0 and making it an integer vector
```{r}
dat$Experience[dat$Experience=="R"] = 0
dat$Experience = as.integer(dat$Experience)
dat$Experience
```

# 4. Performance of players

### Adding an EFF variable to our data frame
Creating Missed_FG variable
```{r}
dat$Missed_FG <- dat$FGA - dat$FGM
dat$Missed_FG
```

Creating Missed_FT variable
```{r}
dat$Missed_FT <- dat$FTA - dat$FTM
dat$Missed_FT
```

Creating PTS variable
```{r}
dat$PTS <- dat$Points3*3 + dat$Points2*2 + dat$FTM
dat$PTS
```

Creating REB variable
```{r}
dat$REB <- dat$DREB + dat$OREB
dat$REB
```

Creating MPG variable
```{r}
dat$MPG <- dat$MIN / dat$GP
dat$MPG
```

Creating EFF variable
```{r}
dat$EFF <- (dat$PTS + dat$REB + dat$AST + dat$STL + dat$BLK - dat$Missed_FG - dat$Missed_FT - dat$TO) / dat$GP
```

Summary statistics of EFF
```{r}
summary(dat$EFF)
ggplot(data = dat, aes(EFF)) + geom_histogram(binwidth = 5)
```

Top-10 players by EFF in decreasing order
```{r}
slice(arrange(dat, desc(EFF)), 1:10)
```

Names of players with negative EFF
```{r}
negEFF <- filter(dat, EFF<0)
negEFF[, "Player"]
```
Correlation between EFF and PTS
```{r}
cor1 <- cor(dat$EFF, dat$PTS)
cor1
```

Correlation between EFF and REB
```{r}
cor2 <- cor(dat$EFF, dat$REB)
cor2
```

Correlation between EFF and AST
```{r}
cor3 <- cor(dat$EFF, dat$AST)
cor3
```

Correlation between EFF and STL
```{r}
cor4 <- cor(dat$EFF, dat$STL)
cor4
```

Correlation between EFF and BLK
```{r}
cor5 <- cor(dat$EFF, dat$BLK)
cor5
```

Correlation between EFF and Missed_FG
```{r}
cor6 <- -cor(dat$EFF, dat$Missed_FG)
cor6
```

Correlation between EFF and Missed_FT
```{r}
cor7 <- -cor(dat$EFF, dat$Missed_FT)
cor7
```

Correlation between EFF and TO
```{r}
cor8 <- -cor(dat$EFF, dat$TO)
cor8
```

```{r}
correlations <- c(cor1, cor2, cor3, cor4, cor5, cor6, cor7, cor8)
cols <- c("coral1", "grey")[(correlations > 0) + 1]
barplot(correlations, main = "Correlations between Player Stats and EFF", names.arg = c("PTS", "REB", "AST", "STL", "BLK", "Missed_FG", "Missed_FT", "TO"), cex.names = 0.4, col = cols)
```


# 5. Efficiency and Salary

```{r}
eff_plot = plot(x = dat$EFF, y = dat$Salary, main="Scatterplot Example", xlab="Efficiency", ylab="Salary")
lines(lowess(dat$EFF, dat$Salary), col=2, lwd=2)

```

Correlation between efficiency and salary
```{r}
cor(dat$EFF, dat$Salary)
```
We can say that efficiency and salary are positively correlated. So the more efficient a player is, the more he is paid.

Creating the player2 data frame
```{r}
player2 <- filter(dat, MPG>20)
player2
```

Creating the player2 scatterplot of efficiency vs salary
```{r}
plot(player2$EFF, player2$Salary)
lines(lowess(player2$EFF, player2$Salary), col=2, lwd=2)
```

Correlation for players who play more than 20 minutes
```{r}
cor(player2$EFF, player2$Salary)
```

Looking at the scatterplot and the correlation coefficient, we can conclude that there is a weaker correlation between efficiency and salary, while it is still positive.

# 6. Comments and Reflection

### 1. What things were hard, even though you saw them in class/lab?
For some reason, making the bar chart was difficult because even though the bars were the correct heights, I had a hard time getting them to display in decreasing order.

### 2. What was easy(-ish) even though we haven't done it in class/lab?
Making the histograms and some of the plots were fairly easy

### 3. Did you need help to complete the assignment? If so, what kind of help?
I needed a little bit of help on question 4. So I ended up working with a friend for that question

### 4. How much time did it take to complete this HW?
It took me about 2 hours

### 5. What was the most time consuming part?
The most time consuming part was the barchart for the correlations for the variables of the EFF formula.

### 6. Was there anything that you did not understand? or fuly grasped?
I am still figuring how to fully use ggplot functions

### 7. Was there anything frustrating in particular?
The barchart was the most frustrating part for me in this assignment and there wasn't much help on Piazza

### 8. Was there anything exciting? Something that you feel proud of?
Not really
