---
title: "hw01-sandeep-tiwari"
author: "Sandeep Tiwari"
date: "9/21/2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# load date file
load("data/nba2017-salary-points.RData")

# list the available objects
ls()
```

### 1. A bid of data preprocessing

## Change Salary from dollars to millions of dollars

```{r}
new_salary <- salary / 1000000
new_salary
```
## Make experience an integer vector
```{r}
experience[experience == "R"] = 0
int_experience <- as.integer(experience)
int_experience
```
## Create more descriptive labels for positions
```{r}
position[position == "PG"] = "point_guard"
position[position == "SG"] = "shoot_guard"
position[position == "SF"] = "small_fwd"
position[position == "PF"] = "power_fwd"
position[position == "C"] = "center"
new_position <- factor(position)
new_position
position_table <- table(new_position)
position_table
```

### 2. Scatterplot of Points and Salary
```{r}
?par()
plot(x = points, y = new_salary, xlab = "Points", ylab = "Salary (in millions)")
```

### 3. Correlation between Points and Salary
```{r}
n <- length(player)
x_mean <- sum(points) / n
y_mean <- sum(new_salary) / n
x_var <- sum((points-x_mean)^2) / (n-1)
y_var <- sum((new_salary-y_mean)^2) / (n-1)
x_sd <- sd(points)
y_sd <- sd(new_salary)
xy_cov <- sum((points-x_mean) * (new_salary-y_mean)) / (n-1)
xy_cor <- xy_cov / (x_sd * y_sd)
```

### 4. Simple Linear Regression
```{r}
slope <- xy_cor * (y_sd / x_sd)
int <- y_mean - (slope * x_mean)
y_hat <- int + (slope * points)
```
# Summary statistics of Y hat
```{r}
summary(y_hat)
```
# Regression Equation
```{r}
y_hat = int + slope * points
```
# Interpretation of slope coefficient b1 
The slope coefficient describes how much the salary changes as the number of points scored changes

# Interpretation of intercept term b0
The intercept term is approximately how much someone earns if they score 0 points

# Predicted salary for a player that scores:
0 points
```{r}
y_hat[points == 0]
```
100 points
```{r}
y_hat[points == 100]
```
500 points
```{r}
int + slope * 500
```
1000 points
```{r}
int + slope * 1000
```
2000 points
```{r}
int + slope * 2000
```

### 5. Plotting the regression line
```{r}
plot1 = plot(x = points, y = new_salary, xlab = "Points", ylab = "Salary (in millions)")
abline(int, slope, lwd = 2, col="blue")
lines(lowess(points, new_salary), col=2, lwd=2)
text(2300, 29, labels = c("lowess"), col="red")
text(2300, 17, labels = c("regression"), col="blue")
```

### 6. Regression residuals and Coefficient of Determination R^2
```{r}
residuals <- new_salary - y_hat
rss <- sum(residuals^2)
tss <- sum((new_salary - y_mean)^2)
determination <- 1 - (rss / tss)
```

### Exploring Position and Experience
```{r}
plot2 = plot(int_experience, new_salary)
lines(lowess(int_experience, new_salary), col=2, lwd=2)
```
## 3D scatterplot
```{r}
#needed to comment this out because was unable to knit the document
#scatterplot3d(x = points, y = int_experience, z = new_salary, xlab="Points", ylab="Years of Experience", zlab="Salary (in Millions)")
```

### Boxplot
```{r}
#?boxplot
boxplot(new_salary~new_position, xlab="Position", ylab="Salary (in millions)")
```
From the scatterplots, it seems as though salary seems to peak after around 5-10 years of experience. Then there seems to be a downwards trend. From the boxplots, the general mean salaries are relatively similar for each position, but the center position has a larger range of salaries, while all the other positions have a few outliers.

### 8. Comments and Reflections
Sometimes figuring out what to input in the different plots to produce the desired graph was difficult, but most everything else was fairly straightforward. For the most part I was able to figure out what to do, but if I needed help, I consulted a friend, TA, or Google. I took about 2 hours to complete this assignment. The most time consuming part was making the graphs. For the most part, I felt like I learned everything this assignment intended to teach.